
\begin{thebibliography}{9}

    \bibitem{GamByDesign}
        G. Zichermann y C. Cunningham,
        {\it Gamification by design: Implementing game mechanics in web and mobile apps.}
        Sebastopol, CA, USA: O'Reilly Media, 2011.
        % http://storage.libre.life/Gamification_by_Design.pdf

    \bibitem{Octalysis}
        Y. Chou,
        {\it Actionable Gamification.}
        Malpitas, CA, USA: Octalysis Media, 2016.
        % Not required, we have this book

    \bibitem{DeterdingDefinition}
        S. Deterding, R. Khaled, L. Nacke y D. Dixon,
        ``Gamification: Toward a definition'',  en
        {\it CHI}, Vancouver, BC, Canada, 2011. p. 2
        % http://gamification-research.org/wp-content/uploads/2011/04/02-Deterding-Khaled-Nacke-Dixon.pdf

    \bibitem{Seaborn}
        K. Seaborn y D.I. Fels
        ``Gamification in theory and action: A survey'',
        {\it International Journal of Human-computer Studies}, vol. 74, no. C, pp. 14-31. Feb. 2015.

    \bibitem{DeterdingGamefulness}
        S. Deterding, D. Dixon, R. Khaled, y L. Nacke.
        $"$From game design elements to gamefulness: defining gamification $"$, en {\it Proc. of the 15th
        International Academic MindTrek Conference: Envisioning Future Media Environments} New York, NY, USA, 2011, pp. 9-15.
        % http://www.rolandhubscher.org/courses/hf765/readings/Deterding_2011.pdf

    \bibitem{Aldemir}
        T. Aldemir, B. Celik y G. Kaplan,
        ``A qualitative investigation of student perceptions of game elements in a gamified course'',
        {\it Comput. Hum. Behav.}, vol. 78, 2018. pp. 235-254.
        % https://daneshyari.com/article/preview/4937006.pdf

    \bibitem{GamInE-Learning}
        Strmečki, D., Bernik, A., Radošević, D. (2015).
        {\it ``Gamification in e-learning: introducing gamified design elements into e-learning systems''}.
        Journal of Computer Science, 11(12), 1108-1117.
        % https://www.researchgate.net/publication/303430512_Gamification_in_E-Learning_Introducing_Gamified_Design_Elements_into_E-Learning_Systems/download

    \bibitem{Lee}
        Lee, J., \& Hammer, J. (2011).
        {\it Gamification in education: What, how, why bother?}
        Academic Exchange Quarterly, 15(2), p. 146.
        % https://www.researchgate.net/publication/258697764_Gamification_in_Education_What_How_Why_Bother

    \bibitem{ForTheWin}
        Werbach, Kevin y Dan Hunter.
        {\it ``For the Win: How Game Thinking Can Revolutionize Your Business.''}
        Harrisburg: Wharton Digital Press, 2012.
        % https://ieeexplore.ieee.org/stamp/stamp.jsp?tp=&arnumber=8586766

    \bibitem{Wood-Reiners}
        Wood, L. C. Reiners, T. (2015).
        {\it ``Gamification''}. In M. Khosrow-Pour (Ed.), Encyclopedia of Information Science and Technology (3rd ed., pp. 3039-3047).
        Hershey, PA: Information Science Reference. DOI:10.4018/978-1-4666-5888-2.ch297
        % https://www.researchgate.net/publication/265337179_Gamification

\end{thebibliography}
